Coexistence with varying environmental filtering intensity

### 2018-06-06 env_intensity.Rmd
SES for NullModel II and plot added.

library(cowplot)
```

@@

# Compute FD metrics (richness, dispersion, eveness)
library(FD)
timestep = 10
FD_prop <- lapply(names(multiple_traits),
                  function(x){
                      patch_sp_timestep <- multiple_traits[[x]]$compo[,,timestep]
                      ab_rel = patch_sp_timestep / rowSums(patch_sp_timestep)
                      ab_rel = ab_rel[!is.na(rowSums(ab_rel)), ]
                      FD_patches <- sapply(seq_len(nrow(ab_rel)), function(y){
                          ab_patch = ab_rel[y, ][ab_rel[y, ] > 0] # Removal of species with null abundance
                          if(length(ab_patch) > 1){ # Patch must contain at least 2 species
                              FD_patch = dbFD(traits[rownames(traits) %in%
                                                         names(ab_patch), ],
                                              a = ab_patch,
                                              w.abun = TRUE)
                              data.frame(prop = x,
                                         patch = rownames(ab_rel[y, , drop = FALSE]),
                                         frich = FD_patch$FRic,
                                         fdis = FD_patch$FDis,
                                         feve = FD_patch$FEve)
                          }
                      })
                      FD_patches <- do.call(rbind, FD_patches)
                  })
FD_prop <- do.call(rbind, FD_prop)

@@

##
# Null models: resample species amongst patches
po = multiple_traits[["prop1"]]$compo[, , 10]
nb_perm = 9#999
# Null model 2 (Bernard-Verdier et al. 2012) => shuffling abundances of species within patches
null_matrices <- replicate(nb_perm, t(apply(po, 1, sample)), simplify = FALSE)
# Convert to relative abundances
null_matrices <- lapply(null_matrices, function(y) y/rowSums(y))
# Add species in column names
null_matrices = lapply(null_matrices, as.data.frame)
null_matrices = lapply(null_matrices, setNames, paste0("species", seq(species)))
# Remove empty rows in null matrices
null_matrices = lapply(null_matrices, function(x) x[!is.na(rowSums(x)), ])

# FD computation over the null matrices
FD_null <- lapply(null_matrices,
                  function(y){
                      FD_null_patches <- sapply(seq_len(nrow(y)), function(z){
                          ab_patch = y[z, ]
                          ab_patch = ab_patch[, colSums(ab_patch) > 0] # Removal of species with null abundance
                          #ab_patch = z[, z > 0, drop = FALSE] 
                          if(length(ab_patch) > 1){ # Patch must contain at least 2 species
                              FD_patch = dbFD(traits[rownames(traits) %in%
                                                         names(ab_patch), ],
                                              a = ab_patch,
                                              w.abun = TRUE)
                              data.frame(prop = "prop1", # => x
                                         patch = rownames(ab_patch),
                                         frich = FD_patch$FRic,
                                         fdis = FD_patch$FDis,
                                         feve = FD_patch$FEve)
                          }
                      })
                      FD_null_patches <- do.call(rbind, FD_null_patches)
                  })
FD_null <- do.call(rbind, FD_null)

library(dplyr)
FD_null2 <- group_by(FD_null, prop)
FD_null2 <- summarise_if(FD_null2,
                         is.numeric,
                         list(mean = function(x) mean(x, na.rm = TRUE),
                              sd = function(x) sd(x, na.rm = TRUE)))

tapply(FD_null$frich, FD_null$patch, mean, na.rm = TRUE)
tapply(FD_null$frich, FD_null$patch, sd, na.rm = TRUE)
FD_SES

# Standard effect size
# (obs - mean(Null)) / sigma(Null)

>>>>

library(cowplot)
library(dplyr)
library(FD)
```

@@

# Compute FD metrics (richness, dispersion, eveness)
timestep = 10
FD_prop <- lapply(names(multiple_traits),
                  function(x){
                      patch_sp_timestep <- multiple_traits[[x]]$compo[,,timestep]
                      ab_rel = patch_sp_timestep / rowSums(patch_sp_timestep)
                      ab_rel = ab_rel[!is.na(rowSums(ab_rel)), ]
                      FD_patches <- sapply(seq_len(nrow(ab_rel)), function(y){
                          ab_patch = ab_rel[y, ][ab_rel[y, ] > 0] # Removal of species with null abundance
                          if(length(ab_patch) > 1){ # Patch must contain at least 2 species
                              FD_patch = dbFD(traits[rownames(traits) %in%
                                                         names(ab_patch), ],
                                              a = ab_patch,
                                              w.abun = TRUE)
                              data.frame(prop = x,
                                         patch = rownames(ab_rel[y, , drop = FALSE]),
                                         frich = FD_patch$FRic,
                                         fdis = FD_patch$FDis,
                                         feve = FD_patch$FEve)
                          }
                      })
                      FD_patches <- do.call(rbind, FD_patches)
                  })
FD_prop <- do.call(rbind, FD_prop)

@@

# Null models: resample species within each patch (NM2)
# FD null for each proportion case
nb_perm = 99#99
timestep = 10
FD_null_prop <- lapply(names(multiple_traits),
                       function(x){
                           # Null model 2 (Bernard-Verdier et al. 2012) => shuffling abundances of species within patches
                           null_matrices <- replicate(nb_perm,
                                                      t(apply(multiple_traits[[x]]$compo[, , timestep], 1, sample)),
                                                      simplify = FALSE)
                           # Convert to relative abundances
                           null_matrices <- lapply(null_matrices, function(y) y/rowSums(y))
                           # Add species in column names
                           null_matrices = lapply(null_matrices, as.data.frame)
                           null_matrices = lapply(null_matrices, setNames, paste0("species", seq(species)))
                           # Remove empty rows in null matrices
                           null_matrices = lapply(null_matrices, function(x) x[!is.na(rowSums(x)), ])

                           # FD computation over the null matrices
                           FD_null <- lapply(null_matrices,
                                             function(y){
                                                 FD_null_patches <- sapply(seq_len(nrow(y)), function(z){
                                                     ab_patch = y[z, ]
                                                     ab_patch = ab_patch[, colSums(ab_patch) > 0] # Removal of species with null abundance
                                                     #ab_patch = z[, z > 0, drop = FALSE] 
                                                     if(length(ab_patch) > 1){ # Patch must contain at least 2 species
                                                         FD_patch = dbFD(traits[rownames(traits) %in%
                                                                                    names(ab_patch), ],
                                                                         a = ab_patch,
                                                                         w.abun = TRUE)
                                                         data.frame(prop = x,
                                                                    patch = rownames(ab_patch),
                                                                    frich = FD_patch$FRic,
                                                                    fdis = FD_patch$FDis,
                                                                    feve = FD_patch$FEve)
                                                     }
                                                 })
                                                 FD_null_patches <- do.call(rbind, FD_null_patches)
                                             })
                           FD_null <- do.call(rbind, FD_null)

                           FD_null <- group_by(FD_null, patch)
                           FD_null <- summarise_if(FD_null,
                                                   is.numeric,
                                                   list(mean = function(x) mean(x, na.rm = TRUE),
                                                        sd = function(x) sd(x, na.rm = TRUE)))
                           FD_null <- data.frame(FD_null)
                           FD_null$prop <- x
                           return(FD_null)
                       })

FD_null_prop <- do.call(rbind, FD_null_prop)

FD_prop <- left_join(FD_prop, FD_null_prop, by = c("prop", "patch"))
FD_prop$frich_SES <- (FD_prop$frich - FD_prop$frich_mean) / FD_prop$frich_sd
FD_prop$fdis_SES <- (FD_prop$fdis - FD_prop$fdis_mean) / FD_prop$fdis_sd
FD_prop$feve_SES <- (FD_prop$feve - FD_prop$feve_mean) / FD_prop$feve_sd

# Plot
plot_prop_frich_SES <- ggplot(FD_prop, aes(1/width, frich_SES, color = as.factor(prop))) +
    geom_line(size = 1) +
    scale_color_manual(values = viridis::viridis(3)) +
    xlab("Environmental intensity (1/width)") +
    ylab("Patch functional richness SES") +
    theme(legend.position = "none")

plot_prop_fdis_SES <- ggplot(FD_prop, aes(1/width, fdis_SES, color = prop)) +
    geom_line(size = 1) +
    scale_color_manual(values = viridis::viridis(3)) +
    xlab("Environmental intensity (1/width)") +
    ylab("Patch functional dispersion SES") +
    theme(legend.position = "none")

plot_prop_feve_SES <- ggplot(FD_prop, aes(1/width, feve_SES, color = prop)) +
    geom_line(size = 1) +
    scale_color_manual("Proportion of traits\ncontributing to R A",
                       values = viridis::viridis(3)) +
    xlab("Environmental intensity (1/width)") +
    ylab("Patch functional evenness SES")

plot_grid(plot_prop_frich_SES, plot_prop_fdis_SES, plot_prop_feve_SES, nrow = 1)
